---
title: "CMP Analysis"
author: "Oahu MPO"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---

```{r setup, include=FALSE, cache = FALSE}
# Note, do not knit cmp_index.Rmd or the cmp project template Rmd files
# manually. Use "R/CMP/creation_script.R"

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
options(scipen=999) # removes sci notation
```

```{r libraries, include = FALSE}
library(tidyverse)
library(htmlTable)
library(DT)
```

# Introduction

The following information presents the results of project analysis performed for
the Congestion Management Program (CMP). Projects are scored based on various
metrics derviced from the model (e.g., change in volume).

The two tables below presents the point categories and possible values. Note
that transit mode share is only applied to transit projects.

```{r roadway scoring rubric}
score_tbl1 <- data_frame(
  `Performance Measure` = c(
    "Change in V/C Ratio",
    "List of Congested Roadways",
    "Transit Mode Share",
    "VMT",
    "VHT",
    "VHD"
  ),
  Increase = c(0, "", 3, 0, 0, 0),
  `No Change` = c(2, "", 1, 1, 1, 1),
  Decrease = c(5, "", 0, 2, 2, 2),
  `New Road` = c(3, "", "", "", "", ""),
  `Cong Road` = c("", 1, "", "", "", "")
)

score_tbl2 <- data_frame(
  `Volume Increase` = c(
    "No Increase",
    "1 - 499",
    "500 - 999",
    "1000 - 1499",
    "1500 - 1999",
    "2000 - 2499",
    "2500 - 2999",
    "3000 - 3499",
    "3500 + "
  ),
  Points = c(0, 1, 2, 3, 4, 5, 6, 7, 8)
)

score_tbl1 %>% htmlTable(
  rnames = FALSE,
  align = "lccccc"
)

score_tbl2 %>% htmlTable(
  rnames = FALSE,
  align = "lc"
)
```


# Project Scores

```{r project scores}
proj_info %>%
  left_join(non_ec, by = c("Model ProjID" = "ProjID")) %>%
  select(-c(rep_link, cong_road)) %>%
  filter(!is.na(score)) %>%
  rename(Score = score) %>%
  arrange(desc(score)) %>%
  kable()
```


# Individual Project Pages
Each link below provides detailed information for a single project.  

```{r, results='asis'}
# The non_ec file is read by creation_script.R
for (id in non_ec$ProjID){
  char_id <- as.character(id)
  cat("<br>[Project ", char_id, "](cmp_proj_", char_id, ".html)", sep = "")
}
```

