---
title: "CMP Project Analysis"
author: "Oahu MPO"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---

```{r, include = FALSE}
# Template for CMP project analysis
```

```{r setup, include=FALSE, cache = FALSE}
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, error=FALSE)
options(scipen=999) # removes sci notation
```

```{r libraries, include = FALSE}
library(tidyverse)
library(maptools)
library(rgdal)
library(leaflet)
library(knitr)
library(htmltools)
library(DT)
```

# Project `r as.character(id)`

## Project Table

```{r}
proj_info %>%
  filter(`Model ProjID` == id) %>%
  kable()
```


## Project Map
In the map below, the blue line shows the extent of the project as modeled.
The link or links highlighted red are the ones selected to report changes in
volume and volume-to-capacity (V/C) ratios.
```{r}
#dir <- paste0("../cmp_proj_", id)
#for testing
dir <- "C:/projects/Honolulu/Version6/OMPORepo/scenarios/CMP_2016/cmp_proj_1"
shp <- paste0(dir, "/reports/proj_shapefile.shp")
  
# shp <- readShapeSpatial(shp, proj4string = CRS("+init=EPSG:4326"))
# shp <- readShapeSpatial(shp, proj4string = CRS("+init=EPSG:4269"))
shp <- readShapeSpatial(shp, proj4string = CRS("+init=nad83:5101"))
shp <- sp::spTransform(shp, CRS("+init=EPSG:4326"))

# Create a shapefile of just the representative links that are used to
# retrieve point data like volume and V/C change
rep_link_shp <- shp[shp$ID %in% rep_links, ]
```

```{r}
leaflet() %>%
  addTiles() %>%
  addPolylines(data = shp) %>%
  addPolylines(data = rep_link_shp, color = "red", opacity = 1)
```

## Project Impact

```{r, include = FALSE}
ec <- read_csv(paste0(ec_dir, "/reports/VMT and Speeds by FT and AT.csv")) %>%
  mutate(scenario = "ec")
proj <- read_csv(paste0(dir, "/reports/VMT and Speeds by FT and AT.csv")) %>%
  mutate(scenario = "project")
```

```{r}
tbl <- bind_rows(ec, proj) %>%
  mutate(AreaType = as.character(AreaType))

class_labels <- c(
  "Freeway",
  "Expressway",
  "Principal Arterial",
  "Minor Arterial",
  "Major Collector",
  "Minor Collector",
  "Local",
  "Ramp",
  "CC", 
  "Total"
)
```

### Overview

```{r}
tbl %>%
  group_by(scenario) %>%
  summarize(VMT = sum(VMT), VHT = sum(VHT), Delay = sum(Delay)) %>%
  gather(key = stat, value = number, -scenario) %>%
  spread(key = scenario, value = number) %>%
  mutate(
    Difference = ec - project,
    `Percent Diff` = round(Difference / ec * 100, 2),
    stat = factor(stat, levels = c("VMT", "VHT", "Delay", ordered = TRUE))
  ) %>%
  arrange(stat) %>%
  rename(Stat = stat, `Baseline / EC` = ec, Project = project) %>%
  kable(
    digits = c(0, 0, 0, 0, 2),
    format.args = list(big.mark = ",")
  )
```

### Vehicle Miles Traveled
Vehicle Miles Traveled (VMT) is a unit that measures travel distance made by a 
vehicle, such as an automobile, van, pickup truck, commercial vehicle, or 
motorcycle. This measure reflects total year 2040 forecasted islandwide VMT 
produced by the CMP project package for all facility and area types for the
two-hour AM peak period. Projects in a CMP package that show a decrease in VMT
receive two points, one point for no change, and zero points for an increase in
VMT.

The table below shows the island-wide change in VMT by facility type and area
type. For scoring, the change in total VMT is used. Because no single project
can have a dramatic influence on island-wide VMT, changes of any amount are
scored as an increase or decrease.
```{r}
# custom container for the following tables
# used to specify "Area Type" colspan
imp_container = htmltools::withTags(table(
  class = 'display',
  thead(
    tr(
      th(colspan = 1, ""),
      th(colspan = 9, "Area Type")
    ),
    tr(
      lapply(c("FClass", "Most Dense", 2:7, "Least Dense", "Total"), th)
    )
  )
))
```

#### With Project
```{r}
vmt_tbl <- tbl %>%
  group_by(AreaType, FClass, scenario) %>%
  summarize(VMT = sum(VMT)) %>%
  spread(key = scenario, value = VMT) %>%
  marg_2d(., "AreaType", "FClass") %>%
  mutate(
    diff = ec - project,
    pct_diff = ifelse(ec == 0, 0, diff / ec)
  ) %>%
  ungroup() %>%
  mutate(FClass = factor(
    FClass, levels = class_labels, ordered = TRUE
  )) %>%
  arrange(FClass)

vmt_tbl %>%
  select(AreaType, FClass, project) %>%
  spread(key = AreaType, value = project) %>%
  datatable(
    container = imp_container, rownames = FALSE,
    extensions = "Buttons",
    options = list(
      dom = "Bt",
      buttons = c("copy", "csv", "pdf")
    )
  ) %>%
  formatCurrency(c(2:10), digits = 0, currency = "") %>%
  formatStyle("Total", fontWeight = "bold") %>%
  formatStyle(
    "FClass", target = "row",
    fontWeight = styleEqual(c("Total"), c("bold"))
  )
```

#### Percent Change from Baseline / EC
```{r}
vmt_tbl %>%
  select(AreaType, FClass, pct_diff) %>%
  spread(key = AreaType, value = pct_diff) %>%
  datatable(
    container = imp_container, rownames = FALSE,
    extensions = "Buttons",
    options = list(
      dom = "Bt",
      buttons = c("copy", "csv", "pdf")
    )
  ) %>%
  formatPercentage(c(2:10), digits = 2) %>%
  formatStyle("Total", fontWeight = "bold") %>%
  formatStyle(
    "FClass", target = "row",
    fontWeight = styleEqual(c("Total"), c("bold"))
  )
```



### Vehicle Hours Traveled
Vehicle Hours Traveled (VHT) is a measurement of the total hours traveled by all
vehicles. It is calculated by multiplying the number of vehicles times the
travel time of those vehicles on a network or specific links within that
network. VHT is an indicator of how additional travel demand influences
congestion in the system from a travel time standpoint. This measure reflects
total year 2040 forecasted islandwide VHT produced by the CMP project package
for all facility and area types for the two-hour AM peak period. Projects in a
CMP package that show a decrease in VHT receive two points, one point for no
change, and zero points for an increase in VHT.

The table below shows the island-wide change in VHT by facility type and area
type. For scoring, the change in total VHT is used. Because no single project
can have a dramatic influence on island-wide VHT, changes of any amount are
scored as an increase or decrease.

#### With Project
```{r}
vht_tbl <- tbl %>%
  group_by(AreaType, FClass, scenario) %>%
  summarize(VHT = sum(VHT)) %>%
  spread(key = scenario, value = VHT) %>%
  marg_2d(., "AreaType", "FClass") %>%
  mutate(
    diff = ec - project,
    pct_diff = ifelse(ec == 0, 0, diff / ec)
  ) %>%
  ungroup() %>%
  mutate(FClass = factor(
    FClass, levels = class_labels, ordered = TRUE
  )) %>%
  arrange(FClass)

vht_tbl %>%
  select(AreaType, FClass, project) %>%
  spread(key = AreaType, value = project) %>%
  datatable(
    container = imp_container, rownames = FALSE,
    extensions = "Buttons",
    options = list(
      dom = "Bt",
      buttons = c("copy", "csv", "pdf")
    )
  ) %>%
  formatCurrency(c(2:10), digits = 0, currency = "") %>%
  formatStyle("Total", fontWeight = "bold") %>%
  formatStyle(
    "FClass", target = "row",
    fontWeight = styleEqual(c("Total"), c("bold"))
  )
```


#### Percent Change from Baseline / EC
```{r}
vht_tbl %>%
  select(AreaType, FClass, pct_diff) %>%
  spread(key = AreaType, value = pct_diff) %>%
  datatable(
    container = imp_container, rownames = FALSE,
    extensions = "Buttons",
    options = list(
      dom = "Bt",
      buttons = c("copy", "csv", "pdf")
    )
  ) %>%
  formatPercentage(c(2:10), digits = 2) %>%
  formatStyle("Total", fontWeight = "bold") %>%
  formatStyle(
    "FClass", target = "row",
    fontWeight = styleEqual(c("Total"), c("bold"))
  )
```


### Delay
Vehicle Hours of Delay (VHD) is defined as the difference between vehicle-hours 
traveled under congested conditions and vehicle hours of travel that would
otherwise be expected under free flow (uncongested) conditions. VHD is
calculated using travel times and travel speeds. This measure reflects total
year 2040 forecasted islandwide VHD produced by the CMP project package for all
facility and area types for the two-hour AM peak period. Projects in a CMP
package that show a decrease in VHD receive two points, one point for no change,
and zero points for an increase in VHD.

The table below shows the island-wide change in VHD by facility type and area
type. For scoring, the change in total VHD is used. Because no single project
can have a dramatic influence on island-wide VHD, changes of any amount are
scored as an increase or decrease.

#### With Project
```{r}
vhd_tbl <- tbl %>%
  group_by(AreaType, FClass, scenario) %>%
  summarize(Delay = sum(Delay)) %>%
  spread(key = scenario, value = Delay) %>%
  marg_2d(., "AreaType", "FClass") %>%
  mutate(
    diff = ec - project,
    pct_diff = ifelse(ec == 0, 0, diff / ec)
  ) %>%
  ungroup() %>%
  mutate(FClass = factor(
    FClass, levels = class_labels, ordered = TRUE
  )) %>%
  arrange(FClass)

vhd_tbl %>%
  select(AreaType, FClass, project) %>%
  spread(key = AreaType, value = project) %>%
  datatable(
    container = imp_container, rownames = FALSE,
    extensions = "Buttons",
    options = list(
      dom = "Bt",
      buttons = c("copy", "csv", "pdf")
    )
  ) %>%
  formatCurrency(c(2:10), digits = 0, currency = "") %>%
  formatStyle("Total", fontWeight = "bold") %>%
  formatStyle(
    "FClass", target = "row",
    fontWeight = styleEqual(c("Total"), c("bold"))
  )
```

#### Percent Change from Baseline / EC
```{r}
vhd_tbl %>%
  select(AreaType, FClass, pct_diff) %>%
  spread(key = AreaType, value = pct_diff) %>%
  datatable(
    container = imp_container, rownames = FALSE,
    extensions = "Buttons",
    options = list(
      dom = "Bt",
      buttons = c("copy", "csv", "pdf")
    )
  ) %>%
  formatPercentage(c(2:10), digits = 2) %>%
  formatStyle("Total", fontWeight = "bold") %>%
  formatStyle(
    "FClass", target = "row",
    fontWeight = styleEqual(c("Total"), c("bold"))
  )
```

## Spot Metrics
For spot measurements, a representative link or links is selected. The
representative links are highlighted red in the project map. Each metric shows
the before and after value, and whether it is considered an increase or a
decrease.

### Spot Volume Change
Vehicle volume is the number of vehicles that use a given roadway over a
specified period of time. Vehicle volumes are presented in concert with V/C
ratio to provide a comparison of project effectiveness. For example, does the
project (or projects) in a CMP project package reduce congestion (reduced V/C
ratio), while carrying more traffic (higher vehicle volumes)? For this
particular metric, points are assigned based on the ability of project to show
higher vehicle volumes on an existing facility. Points range from zero for no
change in comparison to the Baseline alternative during the AM peak period up to
eight points for increases in vehicle volumes between 3,501 to 4,000 vehicle
trips.

The table below shows the change in volume during the AM period. Changes less than 10% are considered no change for scoring.
```{r}
proj_vol <- rep_link_shp@data %>%
  mutate(tot_vol = AB_FLOW_AM + BA_FLOW_AM) %>%
  summarize(sum(tot_vol))

ec_vol <- ec_shp@data %>%
  filter(ID %in% rep_links) %>%
  mutate(tot_vol = AB_FLOW_AM + BA_FLOW_AM) %>%
  summarize(sum(tot_vol))

vol_diff <- proj_vol - ec_vol
  
vol_change <- "No Change"
if (proj_vol > ec_vol * 1.1) vol_change <- "Increase"
if (proj_vol < ec_vol * .9) vol_change <- "Decrease"

proj_vol <- format(round(proj_vol, -1), big.mark = ",")
ec_vol <- format(round(ec_vol, -1), big.mark = ",")

data_frame(
  ` ` = c("Baseline", "With Project", "Change"),
  Value = c(ec_vol, proj_vol, vol_change)
) %>%
  kable(
    format.args = list(big.mark = ",")
  )
```

### Spot V/C
The table below shows the change in maximum volume-to-capacity (V/C) ratios
during the AM period. Changes by +/- .05 are considered no change for scoring.
```{r}
proj_vc <- rep_link_shp@data %>%
  mutate(max_vc = max(AB_VOC_AM + BA_VOC_AM)) %>%
  summarize(max(max_vc))

ec_vc <- ec_shp@data %>%
  filter(ID %in% rep_links) %>%
  mutate(max_vc = max(AB_VOC_AM + BA_VOC_AM)) %>%
  summarize(max(max_vc))

vc_change <- "No Change"
if (proj_vc > ec_vc + .05) {vc_change <- "Increase"}
if (proj_vc < ec_vc - .05) {vc_change <- "Decrease"}

proj_vc <- format(round(proj_vc, 2), big.mark = ",")
ec_vc <- format(round(ec_vc, 2), big.mark = ",")

data_frame(
  ` ` = c("Baseline", "With Project", "Change"),
  Value = c(ec_vc, proj_vc, vc_change)
) %>%
  kable(
    format.args = list(big.mark = ",")
  )
```

## Final Score

```{r}
score <- 0

# v/c score
if (vc_change == "No Change") {score <- score + 2}
if (vc_change == "New Road") {score <- score + 3}
if (vc_change == "Decrease") {score <- score + 5}

# list of congested roadways
if (cong_road == 1) {score <- score + 1}

# Assume no transit mode change for highway projects

# Volume increase
score <- score + case_when(
  vol_diff > 3500 ~ 8,
  vol_diff > 3000 ~ 7,
  vol_diff > 2500 ~ 6,
  vol_diff > 2000 ~ 5,
  vol_diff > 1500 ~ 4,
  vol_diff > 1000 ~ 3,
  vol_diff > 500 ~ 2,
  vol_diff > 0 ~ 1,
  vol_diff <= 0 ~ 0
)

# VMT
vmt_diff <- sum(vmt_tbl$diff)
if (vmt_diff < 0) {score <- score + 2}

# VHT
vht_diff <- sum(vht_tbl$diff)
if (vht_diff < 0) {score <- score + 2}

# VHD
vhd_diff <- sum(vhd_tbl$diff)
if (vhd_diff < 0) {score <- score + 2}
```

<font size = 20>`r score` / 20 points</font>
